Gusto: a toolkit for compatible finite element dynamical cores

### 2024-02-20 source/paper.tex
a bit more text on the software and FML

\def\MM#1{\boldsymbol{#1}}
\newcommand{\pp}[2]{\frac{\partial #1}{\partial #2}}

@@

Prognostic equation sets are each defined in a python class, making
use of the common finite element forms for the transport and
difffusion terms.

** mention physics source terms?

@@

\section{Form Manipulation Labelling}
\label{sec: FML}

\section{Results}
\label{sec: results}
The aim of this section is to present a selection of results that
demonstrate the tools available in Gusto, such as the choice of grids,
order of finite element spaces, different spatial discretisation
options, timestepping schemes and physics. The numerical properties of
these configurations (convergence, stability, accuracy and efficiency)
have either been demonstrated elsewhere, or will be investigated more
thoroughly in future publications.

>>>>

\def\MM#1{\boldsymbol{#1}}
\newcommand{\pp}[2]{\frac{\partial #1}{\partial #2}} 
\newcommand{\JScomment}[1]{\textit{\textbf{Jemma says: #1}}}

@@

\JScomment{Text here depends on what the schematic looks like - this
  paragraph should refer to the schematic and explain all aspects}
Prognostic equation sets are each defined in a python class, making
use of the common finite element forms for the transport and
difffusion terms. The timeloop class contains the definition of the
timestepping procedure and the \texttt{run} method that loops over the
timesteps. It also sets up diagnostic fields and calls the IO as
required. The simplest timeloop class will just apply one time
discretisation method to the equation set but it is common in weather
and climate prediction models to split the different types of terms
and apply a time discretisation separately to each type of term. For
example, the transport and forcing terms might be solved for
separately, such as in **, which enables explicit transport schemes to
be applied to each field. ** something about SIQN structure. Terms
related to the physics are solved for in the appropriate part of the
iteration loop, with physics processes that occur on `fast' timescales
evaluated more frequently and `slower' physics processes evaluated
just once at the end of the timestep. The choice of where to evaluate
the physics processes in the iteration is also based on the
computational cost and the bast way to do this is still an open
research question (**ref).

@@

\section{Form Manipulation Labelling}
\label{sec: FML}
The discussion of the time discretisation so far has been general i.e.
not specific to any particular equation set. This is also true of the
implementation of the timestepping schemes. This enables flexibility
and reproducibility (**elaborate). This functionality is enabled
through form manipulation labelling (FML) which is used to associate a
label to each term in the residual form defined in the equation
class. The timestepping methods are then able to reason about which
manipulations to apply to which forms by applying filters. A simple
example would be a splitting method that evaluates transport terms
separately. The full residual is passed to the transport scheme which
then applies a filter to drop all terms that are not labelled as
transport.

\section{Results}
\label{sec: results}
The aim of this section is to present a selection of results that
demonstrate the tools available in Gusto, such as the choice of grids,
HDiv element families, order of finite element spaces, different
spatial discretisation options, timestepping schemes and physics. The
numerical properties of these configurations (convergence, stability,
accuracy and efficiency) have either been demonstrated elsewhere, or
will be investigated more thoroughly in future publications.
